---
title: "Forecasting Sunspots Activity"
author: "David Jackson"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2) # Includes forecast package
library(tidyverse)
library(janitor)
library(lubridate)
library(tidymodels)
```
## Forecasting Sunspot Actiivity

```{r}
rm(list = ls())
isn <- read_csv2("./kh_spots.csv") %>% 
  remove_empty(which = c("rows","rows")) %>% select(Ymd,g_n,g_s)

```

### Create Time Series

```{r}

```


```{r}
  isn_monthly <- isn %>% group_by(Month = floor_date(Ymd,"month")) %>%
  summarise(G = sum(g_n + g_s)) %>% filter(G >0)
```

### Plot Monthly ISN

```{r}
ggplot(isn_monthly) + geom_col(aes(x=Month,y=G)) + 
  geom_smooth(aes(x=Month,y=G)) + labs(title ="Monthly Sunspot Groups")
```

### Split Data Into Training and Test



```{r}
set.seed(3246)
isn_split <- initial_time_split(isn_monthly, prop = 3/4)
train_isn <- training(isn_split)
test_isn <-  testing(isn_split)
```

#### Plot Training and Test Splits

```{r}
ggplot(train_isn) + geom_line(aes(x=Month,y=G,col="Train Data")) +
  geom_line(data=test_isn,aes(x=Month,y=G,col="Test Data")) +
  labs(title ="Plotting Splits: Training and Test Data")
```

### Forecast Using: meanf, Diff, snaive,

```{r}
Y <- ts(train_isn$G,start=c(1945,4),frequency = 12)
```

#### Mean(f) Forecast

```{r}
mean_fc <- meanf(Y)
autoplot(mean_fc,include=36) + labs(title="Mean Forecast")
```


```{r}
DY <- diff(Y)
ggseasonplot(DY)
```

```{r}
fit <- naive(DY,h=36) # Residual sd: 14.2405
print(fit)
summary(fit)
checkresiduals(fit)
autoplot(fit,include=72)
```


#### seasonal naive model 
 
```{r}
fit <- snaive(DY,h=36) # Residual sd: 68.8911
print(fit)
summary(fit)
checkresiduals(fit)
autoplot(fit,include=60)
```

### ETS Forecast

```{r}
fit_ets <- ets(Y) # sigma:  218.8133
summary(fit_ets)
checkresiduals(fit_ets)
autoplot(fit_ets)
```

### auto.arima model
#### Best fit: Best model: ARIMA(0,1,1)(2,1,2)[12]

```{r}
fit_arima <- auto.arima(Y,d=1,D=1,stepwise = FALSE,approximation = FALSE, trace= TRUE)
summary(fit_arima)
checkresiduals(fit_arima)
autoplot(fit_arima)
# sqrt() # SD 197.8105
```

#### Forecast using Arima

```{r}
fcast1 <- forecast(fit_arima,h=24)
autoplot(fcast1)
autoplot(fcast1,include=365) + labs(title ="365 Year")
autoplot(fcast1,include=60) + labs(title = "Last 5 Years") 
```

### Forecast From 1945 - today - 1 day.
```{r}
X <- ts(isn_monthly$G,start=c(1945,4),frequency = 12)
XX <- diff(X)
```

```{r}
fit <- snaive(XX,h=36) # 
print(fit)
summary(fit)
checkresiduals(fit)

autoplot(fit) + labs(title ="Compete Data")

autoplot(fit,include=36) + labs(title ="Include 36")
autoplot(fit,include=60) + labs(title ="5 Years")
autoplot(fit,include=365) + labs(title ="includes 365")

```


### auto.arima model: Entire Data
#### Best fit: Best model: ARIMA(0,1,1)(2,1,2)[12]

```{r}
fit_arima <- auto.arima(X,d=1,D=1,stepwise = FALSE,approximation = FALSE, trace= TRUE)
summary(fit_arima)
checkresiduals(fit_arima)
autoplot(fit_arima)
# sqrt() # SD 197.8105
```

#### Forecast using Arima: entire data

```{r}
fcast1 <- forecast(fit_arima,h=36)
autoplot(fcast1) + labs(title = "Full Forecast")
autoplot(fcast1,include=365) + labs(title ="Include 365")
autoplot(fcast1,include=180) + labs(title ="Include 180")
autoplot(fcast1,include=60) + labs(title = "Include 60") 
```
